---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# acresFSA

<!-- badges: start -->
<!-- badges: end -->

The acresFSA R-package contains a selection of data downloaded from the United States Department of Agriculture (USDA) Farm Service Agency (FSA) website.

> "Farm Service Agency policy requires that producers participating in several programs submit an annual report regarding all cropland use on their farms. These programs include Agriculture Risk Coverage (ARC) and Price Loss Coverage (PLC). Reporting also applies to those who receive marketing assistance loans or loan deficiency payments. Failure to file an accurate and timely acreage report for all crops and land uses can result in loss of program benefits. Producers are required to self report all cropland on each farm to FSA annually. FSA uses these data to determine payment eligibility (land must be in an eligible agricultural use to qualify for payments) and to calculate losses for various disaster programs. Data are reported in the following categories: planted; prevented planted; and failed. In addition, the National Agricultural Statistics Service uses FSA planted acreage data to complement their survey data."

<https://www.fsa.usda.gov/news-room/efoia/electronic-reading-room/frequently-requested-information/crop-acreage-data/index>

# Getting Started
The [Getting Started](https://capellett.github.io/acresFSA/articles/acresfsa.html) article is intended to be accessible to the general public. It includes a download link and an interactive data exploration tool.

# For R Users
The acresFSA R package exports a single data object, named `acresFSA`. The package also contains articles (AKA vignettes), plotting functions, and a few custom helper functions. The vignettes document the methods used to compile and clean up the data.

You can install the development version of acresFSA from [GitHub](https://github.com/) with:

```{r eval=F}
# install.packages("pak")
pak::pak("capellett/acresFSA")
```

